---
title: "Homework # 1"
author: "Holly Finertie (HF2379), Due:"
date: 2019-09-20
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem1

## Create a dataframe:

```{r problem1_dataframe}
set.seed(1)

prob1_df = tibble(
  prob1_sample = rnorm(8, mean = 0, sd = 1),
  prob1_logic_vec = prob1_sample > 0,
  prob1_char_vec = c("I","Am", "A", "Little", "Confused", "Is", "This", "Right"),
  prob1_factor_vec = factor(c("yes", "no", "maybe", "yes", "no", "maybe", "yes", "no"))
)
```

## Mean of each variable:

* **The mean of the sample size is `r mean(pull(prob1_df, prob1_sample))`**. We are able to take the mean because the sample size is a numeric data type. 

* **The mean of the logical statement is `r mean(pull(prob1_df, prob1_logic_vec))`**. We are able to take the mean of the logical statement because it is a Boolean data type (TRUE=1 and FALSE=0). Even if the output is displayed as TRUE/FALSE, the values of 1 and 0 are in the background. 

* **The mean of the character vector is `r mean(pull(prob1_df, prob1_char_vec))`**. We are not able to take the mean of the character vector because it is text/character data type that does not correspond with numeric values. The error states that we can only take the mean of a numeric or logical argument.

* **The mean of the factor vector is `r mean(pull(prob1_df, prob1_factor_vec))`**. We are unable to take the mean of the factor vector because it is a text/character data type that does not correspond with numeric values. The error states that we can only take the mean of a numeric or logical argument.


## First Code Chunk: convert to numeric


```{r problem1_convert, eval=FALSE}
as.numeric(pull(prob1_df, prob1_logic_vec))
as.numeric(pull(prob1_df, prob1_char_vec))
as.numeric(pull(prob1_df, prob1_factor_vec))
```

The as.numeric function works on the logical and factor vectors, but not the character vector.

**Logical**: Because the logical vector is a Boolean data type, we can convert the TRUE / FALSE results to 1 / 0 respectively.  

**Character**: I got the error "NAs introduced by coercion" showing that the as.numeric function did not work on the character vector.  

**Factor**: The as.numeric function works on the factor vector even when we were unable to take the mean. Becuase there were 3 levels, the as.numeric function converted the different levels to 1, 2, and 3. 


## Second Code Chunk: convert and multiply

```{r problem1_secondchunk}
as.numeric(pull(prob1_df, prob1_logic_vec))*pull(prob1_df, prob1_sample)
as.factor(pull(prob1_df, prob1_logic_vec))*pull(prob1_df, prob1_sample)
as.numeric(as.factor(pull(prob1_df, prob1_logic_vec)))*pull(prob1_df, prob1_sample)
```

**Numeric**: We were able to convert the logical vector to numeric vector and multiply by our random sample.

**Factor**: We were unable to multiply the converted logical vector. We got the error: '*' not meaningful for factors. 

**To Factor, to numeric**: We were able to convert the logical vector into factor and then into numeric and multiply by our random sample. 


# Problem 2

```{r problem2_dataframe}
set.seed(2)

prob2_df = tibble(
  x = rnorm(500, mean = 0, sd = 1),
  y = rnorm(500, mean = 0, sd = 1),
  prob2_logic_vec = (x + y) > 1,
  prob2_num_vec = as.numeric(prob2_logic_vec),
  prob2_factor_vec = as.factor(prob2_logic_vec)
)
```

## Describe new data frame

* This dataset contains `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns.

* The mean of x is `r mean(pull(prob2_df, x))`.

* The median of x is `r median(pull(prob2_df, x))`.

* The standard deviation of x is `r sd(pull(prob2_df, x))`.

* The proportion of cases for which x + y > 1 is `r mean(pull(prob2_df, prob2_num_vec))`.

## Plots

```{r problem2_plot_logical}
plot1 = ggplot(prob2_df, aes(x = x, y = y, color = prob2_logic_vec)) + geom_point()
plot1
```

* This plot has two colors, red and blue, because we set the colors to equal the logical vector which was two outcomes: FALSE and TRUE. 


```{r problem2_plot_nvector}
plot2 = ggplot(prob2_df, aes(x = x, y = y, color = prob2_num_vec)) + geom_point()
plot2
```


* This plot's colors range from light blue (1.0) to dark blue (0.0) because we set the color equal to the numeric vector which was outcomes 0 and 1. This numeric vector was created from the logical vector (TRUE=1 and FALSE=0). While the color legend provides a range of colors from 0-1, the output only shows dots that correspond to 0 and 1 identical to plot 1 above. 


```{r problem2_plot_fvector}
plot3 = ggplot(prob2_df, aes(x = x, y = y, color = prob2_factor_vec)) + geom_point()
plot3
```


* This plot has two colors, red and blue, because we set the colors equal to the factor vector which was two outcomes: FALSE and TRUE. This factor vector was created from the logical vector and therefore the graph is identical to plot 1. 


## Saving my first scatterplot

```{r}
ggsave(plot1, file = "plot1.png")
```

Done with HW 1! 


